---
title: "dplyr"
output: html_document
---

##Tiffany Gonzalez

Junior at Mercyhurst University. 
MS in Data Science | BA in Accounting 

#Filtering Columns and Rows

When using dplyr, you can filter columns and rows by using select and filter. Let's look at an example using the Lahman baseball database. We first have to load the Lahman and dplyr packages.

```{r message=FALSE, warning =FALSE}
library(Lahman)
library(dplyr)
```

To see the homerun totals for the 1927 New York Yankee we can run the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(teamID=='NYA' & yearID==1927)
```

Now, if we would like to see the instances where any player had more than 40 HR but less than 60 strikeouts (SO), we could run the following code:

```{r}
Batting%>%
  select(playerID,yearID,teamID,HR,SO)%>%
  filter(HR>40 & SO<60)
```

Next, if you're wondering how many Phillies players in the 1970's hit more than 30 HR's, we could run this code:
```{r message=FALSE, warning=FALSE}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>30 & teamID=='PHI' & yearID>=1970 & yearID<=1979)
```

To find every instance of a player hitting more than 50 HR's we could run the following code... However let's have the players with the most HR's at the top!
```{r message=FALSE, warning=FALSE}
Batting%>%
  select(playerID,yearID,teamID,HR)%>%
  filter(HR>50)%>%
  arrange(desc(HR))
```

Lastly, to find a player who has had at least 400 at-bats (AB), but has less than 10 strike outs strikes, we could run the following code... However let's have players with the smallest strikeout number at the top!
```{r message=FALSE, warning=FALSE}
Batting%>%
  select(playerID,yearID,teamID,AB,SO)%>%
  filter(AB>=400 & SO<10)%>%
  arrange(SO)
